---
title: "Cylinder"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Introduction

Input: 87 trials with RSD=0.4, row of 10 plants. We extracted flat section on the top of tube. We extracted points at most 0.5 cm above the top and at least below the top and with y coordinate between -0.5 and 0.5. This means we extracted 1cm wide flat strip of height 1cm. The total volume of seen dirt is thus 1x1xlength=150cm^3.

## Data

For each insertion angle of tube, we saved a .csv file:
data_cylinder1-2_top30/cylinder1-2_top30_424582000.csv

data files: [depth, numPoints,  totalLength, surfaceArea, volume, radiiAvg, radiiSD]

### Extract Data

```{r,  results="hide"}
angles <- c(30) #seq(30,90, by=5)

# df <- data.frame(matrix(ncol = 3, nrow = 0))
# x <- c("observed", "actual", "angle")
# colnames(df) <- x
observed <- NULL
true <- NULL
df <- NULL

mydir <- file.path(Sys.getenv('HOMEPATH'), paste0("Box Sync/ROOTS/models/Rootbox_v6b/ROGER/Trial Data/Trial_Data_012/"))

for(angle in angles) {
  datadir <- file.path(mydir, paste0("data_cylinder1-2_top", angle))
  # Get the files names
  files <- list.files(path = datadir, pattern="*.csv", full.names = TRUE)

  # Get observed data

  # First apply read.csv, then rbind
  observed_list <- lapply(files,   
                          function(x) {
                            cbind(read.csv(x, stringsAsFactors = FALSE, header=FALSE), file = basename(x))
                          })
  observed <- do.call(rbind, observed_list)

  # get true volumes
  truth_dir <- file.path(mydir, "/data_all_csv")
  truth_files <- list.files(path = truth_dir, pattern = "*.csv", full.names = TRUE)
  # First apply read.csv, then rbind
  truth_list <- lapply(truth_files, 
                       function(x){
                         read.csv(x, stringsAsFactors = FALSE, header=FALSE)
                       } )
  truth <- do.call(rbind, truth_list)
  
  df <- rbind(df, cbind(observed, truth[-1], rep(angle, times = nrow(observed))))
}

colnames(df) <- c("depth", "numPoints",  "totalLength", "surfaceArea", "volume", "radiiAvg", "radiiSD",  "file", "numPoints_T",  "totalLength_T", "surfaceArea_T", "volume_T", "radiiAvg_T", "radiiSD_T", "angle")

df$angle <- factor(df$angle)
df$depth_levels <- cut(df$depth, breaks=15)
df$depth_levels <- factor(df$depth_levels)
```


```{r,  results="hide"}
# install.packages(pkgs = c('Hmisc', 'psych', 'car', 'ggplot2', 'rgl', 'lazyeval'), dependencies = TRUE)
# install.packages("Hmisc", dependencies = TRUE)
# install.packages("psych", dependencies = TRUE)
# install.packages("car", dependencies = TRUE)
# install.packages("ggplot2", dependencies = TRUE)
# install.packages("rgl", dependencies = TRUE)
# install.packages("lazyeval", lib = "C:/Users/sapetka/Documents/R/win-library/3.4")

library(Hmisc)
library(psych)
library(car)
library(ggplot2)
library(rgl)
```

```{r, echo=TRUE}
vol_by_trial <- by(df$volume, df$file, sum)
vol_by_trial_T <- by(df$volume_T, df$file, sum)
mean(vol_by_trial_T / vol_by_trial)
plot(vol_by_trial, vol_by_trial_T)
```

```{r, echo=TRUE}
fit <- lm(volume_T~volume+depth, data=df)
summary(fit)
vif(fit)
```

### Select variables for regression modelling
```{r, echo=TRUE}
data <- subset(df, select = c(depth, volume, volume_T))
pairs.panels(data, col="red")
```

### Split data into training and validation samples
```{r, echo=TRUE}

require(caTools)
train <- data[1:6450,]
valid <- data[6451:12900,]
```

### Create a model which avoids multi-collinearity and checks the p-values
```{r, echo=TRUE}
fit<-lm(volume_T~volume+depth, data=train)
summary(fit)
vif(fit)
```

# Yeo-Johnson Power Transformation (on data including zeros)
### Transform non-linear vars
```{r, echo=TRUE}
#this only works on positive values, but we have many zeros
library(MASS)

depth<-train$depth
volume<-train$volume
volume_T<-train$volume_T

lambda <- boxCox(volume_T~volume+depth, family="yjPower", plotit = TRUE)  
lambda<-lambda$x[which(lambda$y==max(lambda$y))]   #lambda=0.2222222

lambda1 <- boxCox(volume_T~1, family="yjPower", plotit = TRUE)  
lambda1<-lambda1$x[which(lambda1$y==max(lambda1$y))]

lambda2 <- boxCox(volume~1, family="yjPower", plotit = TRUE)  
lambda2<-lambda2$x[which(lambda2$y==max(lambda2$y))]

#yjPower https://rweb.stat.umn.edu/R/site-library/car/html/bcPower.html
# yjPower(X)=((X+1)^(lambda)-1)/lambda
volume_yj <- yjPower(volume, lambda=lambda)
volume_T_yj <- yjPower(volume_T, lambda=lambda)
train_yj<-data.frame(depth=depth, volume=volume_yj, volume_T=volume_T_yj)
valid_yj<-data.frame(depth=valid$depth, volume=yjPower(valid$volume, lambda=lambda), volume_T=yjPower(valid$volume_T, lambda=lambda))
pairs.panels(train_yj, col="red")

fit_yj<-lm(volume_T~volume+depth, data=train_yj)
plot(fit_yj)
summary(fit_yj) 
```

### Check for non-linearity property
```{r, echo=TRUE}
crPlots(fit_yj)
```

#### Now explore the model and results (we will be using data)
```{r, echo=TRUE}
crPlots(fit_yj)
pairs.panels(train_yj, col="red")
```


###predict the outcome of the testing data
```{r, echo=TRUE}
predicted_yj <- predict(fit_yj, newdata=valid_yj[ ,-3], se.fit=FALSE)
```

### what is the proportion variation explained in the outcome of the testing data?
### i.e., what is 1-(SSerror/SStotal)
```{r, echo=TRUE}
actual_yj <- yjPower(valid$volume_T, lambda=lambda) 
rsq <- 1-sum((actual_yj-predicted_yj)^2)/sum((actual_yj-mean(actual_yj))^2)
print(rsq)
```

```{r, echo=TRUE}
vstats <- summary(train_yj$volume)
vstats

vMin = as.numeric(vstats[1]) # Just want numbers not data frames
v1Q = as.numeric(vstats[2])
vMedian = as.numeric(vstats[3])
v3Q = as.numeric(vstats[5])
vMax = as.numeric(vstats[6])
```

### Next: Consider glm and gls, especially for logistic regression


### Add regression lines to the plot
```{r, echo=TRUE}
panel.lm = function(x, y, ...) {
  tmp <- lm(y ~ x, na.action = na.omit)
  if(length(x)>4) {
    points(x, y, ...)
    abline(tmp, col="blue")
    panel.smooth(x, y, col.smooth="red", ...)
  } else {
    points(x, y, ...)
  }
}
```

```{r, echo=TRUE}
coplot(volume_T ~ depth | volume, data=train_yj, 
       col="orange", panel=panel.lm)
```

##### multiple regression 3D plots

### Let us check the impact of variable selection on the model
```{r, echo=TRUE}
scatter3d(z=train_yj$depth, x=train_yj$volume, y=train_yj$volume_T)
```


# Remove zeros, then log transform
```{r, echo=TRUE}
#this only works on positive values, but we have many zeros
library(MASS)
train <- data[1:6450,]
valid <- data[6451:12900,]

depth<-train$depth[train$volume!=0]
volume<-train$volume[train$volume!=0]
volume_T<-train$volume_T[train$volume!=0]
train_log<-data.frame(depth, volume=log(volume), volume_T=log(volume_T))

valid_log<-data.frame(depth=valid$depth[valid$volume!=0], volume=log(valid$volume[valid$volume!=0]), volume_T=log(valid$volume_T[valid$volume!=0]))

pairs.panels(train_log, col="red")

fit_log<-lm(volume_T~depth+volume, data=train_log)
plot(fit_log)
summary(fit_log) 
crPlots(fit_log)
```




```{r, echo=TRUE}
fraction_of_zeros <- function(x) {
  sum(x==0)/length(x)
}
train$p0<-NULL
train$p0<-ave(train$volume, train$depth, FUN= fraction_of_zeros)
plot(train$p0,train$depth)

train$p0_T<-NULL
train$p0_T<-ave(train$volume_T, train$depth, FUN=fraction_of_zeros)
plot(train$p0_T,train$depth)

valid$p0<-ave(valid$volume, valid$depth, FUN= fraction_of_zeros)
valid$p0_T<-ave(valid$volume_T, valid$depth, FUN=fraction_of_zeros)

```

```{r, echo=TRUE}
pairs.panels(train_log, col="red")
fit_log<-lm(volume_T~depth+volume, data=train_log)
plot(fit_log)
summary(fit_log) 
```


###predict the outcome of the testing data
```{r, echo=TRUE}

valid$logVol_pred[valid$volume!=0] <-predict(fit_log, newdata=valid_log[ ,-3], se.fit=FALSE)

# predicted volume for non-zero observed volume, equals predicted$volume_mean(1-p0)
valid$logVol_mean <- ave(valid$logVol_pred, valid$depth, FUN= function(x) mean(x,na.rm=TRUE))
valid$logVol_mean[is.na(valid$logVol_mean)] <-0 

valid$logVol_pred[valid$volume==0] <- valid$logVol_mean[valid$volume==0]*(1-valid$p0[valid$volume==0])
#valid$logVol_pred[valid$volume!=0] <- valid$logVol_pred[valid$volume!=0] #*(1-valid$p0[valid$volume==0])

nonzero <- valid[valid$logVol_pred!=0,]
nonzero$vol_pred<-exp(nonzero$logVol_pred)

plot(nonzero$vol_pred, nonzero$volume_T)

```

### what is the proportion variation explained in the outcome of the testing data?
### i.e., what is 1-(SSerror/SStotal)
```{r, echo=TRUE}
actual <- valid$volume_T
predicted<-exp(valid$logVol_pred)
rsq <- 1-sum((actual-predicted)^2)/sum((actual-mean(actual))^2)
print(rsq)
```

### Analysis depth versus observed volume

```{r, echo=TRUE}
means<- tapply(df$volume, df$depth_levels, mean)
variance <- tapply(df$volume, df$depth_levels, var)

plot(means)
boxplot(df$volume ~ df$depth_levels, xlab="depth", ylab="observed volume", col=rainbow(13))

aov<- aov(df$volume ~ df$depth_levels)
summary(aov)
```

### Adding column residuals to df

```{r, echo=TRUE}
residuals<-NULL
fit<-NULL
angles_30<-c(45)
for (angle in angles) {
  d<-df[df$angle==angle,]
  lm <- lm(d$volume_T ~ d$volume)
  fit <- cbind(fit, lm)
  #s<- summary(fit)
  residuals<-c(residuals, abs(residuals(lm)))
}

plot(d$volume, d$volume_T)
abline(lm)

df$residuals <- residuals
plot(rev(unique(df$depth)), tapply(df$residuals, df$depth, mean))

```

## Fit for the tube (angle 30? with the ground)

### Linear fit true_volume ~ observed_volume

```{r, echo=TRUE}
d<-df[df$angle==30,]
x<-d$volume
y<-d$volume_T
lmXY<-lm(y~x)
plot(x,y)
abline(lmXY)
```

###  Linear fit true_volume ~ observed_volume*ext(depth)
```{r,echo=TRUE}
x<-df$volume
y<-df$volume_T
z<-df$depth
lmXYZ<-lm(y~1+x+z+x*z)
summary(lmXYZ)
plot(lmXYZ)
```

### Linear fit log(true_volume) ~ log(observed_volume)*ext(depth)
```{r,echo=TRUE}
x<-df$volume
y<-df$volume_T
z<-df$depth
loglmXYZ<-lm(log(y+0.0001)~1+log(x+0.0001)+z+log(x+0.0001)*z)
summary(loglmXYZ)
plot(lmXYZ)
```

### Comparison anova
```{r,echo=TRUE}

# the smaller the AIC or BIC, the better the fit.
AIC(lmXY,lmXYZ,loglmXYZ)
```


###  Linear fit observed_volume ~ ext(depth)
```{r,echo=TRUE}
lmXZ<-lm(log(x+0.0001)~z)
summary(lmXZ)
```

###  Linear fit true_volume ~ ext(depth)
```{r,echo=TRUE}
lmYZ<-lm(log(y+0.0001)~z)
summary(lmYZ)
```

