---
title: "helix"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Introduction

Here we analyze performance of helices.

Input: 87 trials with RSD=0.4, row of 10 plants. We extracted flat section on the top of tube. 

## Data

For each insertion angle of helix, we saved a .csv file:
data_helix_A10_alpha30_stretch1/424582000.csv

data files: [depth, numPoints,  totalLength, surfaceArea, volume, radiiAvg, radiiSD]

### Extract Data

```{r}

df<-NULL
observed<-NULL
true<-NULL

  # Get observed data
  setwd("C:/Users/sapetka/Box Sync/ROOTS/models/Rootbox_v6b/ROGER/Trial Data/Trial_Data_012/data_helix_A10_alpha30_stretch1")
  files <- list.files(pattern="*.txt")
  # First apply read.table, then rbind
  observed = do.call(rbind, lapply(files, function(x) {
    ret<-read.table(x, sep="")
    ret$file <- x 
    ret
    }
  ))
  
  # get true volumes
  setwd("C:/Users/sapetka/Box Sync/ROOTS/models/Rootbox_v6b/ROGER/Trial Data/Trial_Data_012/data_all_csv")
  files = list.files(pattern="*.csv")
  # First apply read.csv, then rbind
  true = do.call(rbind, lapply(files,  function(x) {read.csv(x, stringsAsFactors = FALSE, header=FALSE)}))
  
  df <- rbind(df,cbind(observed,true[-1]))

colnames(df) <- c("depth", "numPoints",  "totalLength", "surfaceArea", "volume", "radiiAvg", "radiiSD", "file", "numPoints_T",  "totalLength_T", "surfaceArea_T", "volume_T", "radiiAvg_T", "radiiSD_T")

df$depth_levels<-cut(df$depth,breaks=15)
df$depth_levels <- factor(df$depth_levels)
```


```{r}
# install.packages("Hmisc", dependencies = TRUE)
# install.packages("psych", dependencies = TRUE)
# install.packages("car", dependencies = TRUE)
# install.packages("ggplot2", dependencies = TRUE)
# install.packages("rgl", dependencies = TRUE)
# install.packages("lazyeval", lib = "C:/Users/sapetka/Documents/R/win-library/3.4")

library(Hmisc)
library(psych)
library(car)
library(ggplot2)
library(rgl)
```


```{r, echo=TRUE}
vol_by_trial<-by(df$volume, df$file, sum)
vol_by_trial_T<-by(df$volume_T, df$file, sum)
mean(vol_by_trial_T/vol_by_trial)
plot(vol_by_trial, vol_by_trial_T)
```


## This is now incorrect. Goal: Obtain RMSE of total volume at each level.Do the same for tube. Then compare.
```{r, echo=TRUE}
depth<-df$depth[1:150]
RMSE<-NULL
for(d in depth) {
df$volume[df$depth==d]
df$volume_T[df$depth==d]  

res<-lm(df$volume[df$depth==d]~df$volume_T[df$depth==d])
RSS <- c(crossprod(res$residuals))
#Mean squared error:
MSE <- RSS / length(res$residuals)
#append RMSE:
RMSE <- append(RMSE,sqrt(MSE))
}

plot(depth,RMSE)

plot(by(df$volume_T,df$depth, sum))
```

```{r, echo=TRUE}
fit<-lm(volume_T~volume+depth, data=df)
summary(fit)
vif(fit)
```

### Select variables for regression modelling
```{r, echo=TRUE}
data <- subset(df, select = c(depth, volume, volume_T))
pairs.panels(data, col="red")
```

### Split data into training and validation samples
```{r, echo=TRUE}

require(caTools)
train <- data[1:6450,]
valid <- data[6451:12900,]
```

### Create a model which avoids multi-collinearity and checks the p-values
```{r, echo=TRUE}
fit<-lm(volume_T~volume+depth, data=train)
summary(fit)
vif(fit)
```


# Yeo-Johnson Power Transformation (on data including zeros)
```{r, echo=TRUE}
#this only works on positive values, but we have many zeros
library(MASS)

depth<-train$depth
volume<-train$volume
volume_T<-train$volume_T

lambda <- boxCox(volume_T~volume+depth, family="yjPower", plotit = TRUE)  
lambda<-lambda$x[which(lambda$y==max(lambda$y))]   #lambda=0.2222222

lambda1 <- boxCox(volume_T~1, family="yjPower", plotit = TRUE)  
lambda1<-lambda1$x[which(lambda1$y==max(lambda1$y))]

lambda2 <- boxCox(volume~1, family="yjPower", plotit = TRUE)  
lambda2<-lambda2$x[which(lambda2$y==max(lambda2$y))]

#yjPower https://rweb.stat.umn.edu/R/site-library/car/html/bcPower.html
# yjPower(X)=((X+1)^(lambda)-1)/lambda
# maybe better to use lambda1 for volume and lambda2 for volume_T
volume_yj <- yjPower(volume, lambda=lambda)
volume_T_yj <- yjPower(volume_T, lambda=lambda)
train_yj<-data.frame(depth=depth, volume=volume_yj, volume_T=volume_T_yj)
valid_yj<-data.frame(depth=valid$depth, volume=yjPower(valid$volume, lambda=lambda), volume_T=yjPower(valid$volume_T, lambda=lambda))
pairs.panels(train_yj, col="red")

fit_yj<-lm(volume_T~volume+depth, data=train_yj)
plot(fit_yj)
summary(fit_yj) 
```

### Check for non-linearity property
```{r, echo=TRUE}
crPlots(fit_yj)
```


#### Explore the model and results
```{r, echo=TRUE}
crPlots(fit_yj)
pairs.panels(train_yj, col="red")
```



###predict the outcome of the testing data
```{r, echo=TRUE}
predicted_yj <- predict(fit_yj, newdata=valid_yj[ ,-3], se.fit=FALSE)
```

what is the proportion variation explained in the outcome of the testing data?
i.e., what is 1-(SSerror/SStotal)
```{r, echo=TRUE}
actual_yj <- yjPower(valid$volume_T, lambda=lambda) 
rsq <- 1-sum((actual_yj-predicted_yj)^2)/sum((actual_yj-mean(actual_yj))^2)
print(rsq)
```

```{r, echo=TRUE}
vstats <- summary(train_yj$volume)
vstats

vMin = as.numeric(vstats[1]) # Just want numbers not data frames
v1Q = as.numeric(vstats[2])
vMedian = as.numeric(vstats[3])
v3Q = as.numeric(vstats[5])
vMax = as.numeric(vstats[6])
```

# Consider glm and gls, especially for logistic regression

Coplots:  Plot with depth variable as conditioning var
    We plot with depth in different ranges and multiple plots.
 Add regression lines to the plot
```{r, echo=TRUE}
panel.lm = function(x, y, ...) {
  tmp <- lm(y ~ x, na.action = na.omit)
  if(length(x)>4) {
    points(x, y, ...)
    abline(tmp, col="blue")
    panel.smooth(x, y, col.smooth="red", ...)
  } else {
    points(x, y, ...)
  }
}
```

### Plot with panels featuring regression and loess lines
```{r, echo=TRUE}
coplot(volume_T ~ depth | volume, data=train_yj, 
       col="orange", panel=panel.lm)
```

### multiple regression 3D plots
```{r, echo=TRUE}
scatter3d(z=train_yj$depth, x=train_yj$volume, y=train_yj$volume_T)
```


# Remove zeros, then log transform
```{r, echo=TRUE}
#this only works on positive values, but we have many zeros
library(MASS)
train <- data[1:6450,]
valid <- data[6451:12900,]

depth<-train$depth[train$volume!=0]
volume<-train$volume[train$volume!=0]
volume_T<-train$volume_T[train$volume!=0]
train_log<-data.frame(depth, volume=log(volume), volume_T=log(volume_T))

valid_log<-data.frame(depth=valid$depth[valid$volume!=0], volume=log(valid$volume[valid$volume!=0]), volume_T=log(valid$volume_T[valid$volume!=0]))

pairs.panels(train_log, col="red")

fit_log<-lm(volume_T~depth+volume, data=train_log)
plot(fit_log)
summary(fit_log) 
crPlots(fit_log)
```


```{r, echo=TRUE}
fraction_of_zeros <- function(x) {
  sum(x==0)/length(x)
}
train$p0<-NULL
train$p0<-ave(train$volume, train$depth, FUN= fraction_of_zeros)
plot(train$p0,train$depth)

train$p0_T<-NULL
train$p0_T<-ave(train$volume_T, train$depth, FUN=fraction_of_zeros)
plot(train$p0_T,train$depth)

valid$p0<-ave(valid$volume, valid$depth, FUN= fraction_of_zeros)
valid$p0_T<-ave(valid$volume_T, valid$depth, FUN=fraction_of_zeros)

```

```{r, echo=TRUE}
pairs.panels(train_log, col="red")
fit_log<-lm(volume_T~depth+volume, data=train_log)
plot(fit_log)
summary(fit_log) 
```


###predict the outcome of the testing data
```{r, echo=TRUE}

valid$logVol_pred[valid$volume!=0] <-predict(fit_log, newdata=valid_log[ ,-3], se.fit=FALSE)

# predicted volume for non-zero observed volume, equals predicted$volume_mean(1-p0)
valid$logVol_mean <- ave(valid$logVol_pred, valid$depth, FUN= function(x) mean(x,na.rm=TRUE))
valid$logVol_mean[is.na(valid$logVol_mean)] <-0 

valid$logVol_pred[valid$volume==0] <- valid$logVol_mean[valid$volume==0]*(1-valid$p0[valid$volume==0])
#valid$logVol_pred[valid$volume!=0] <- valid$logVol_pred[valid$volume!=0] #*(1-valid$p0[valid$volume==0])

nonzero <- valid[valid$logVol_pred!=0,]
nonzero$vol_pred<-exp(nonzero$logVol_pred)

plot(nonzero$vol_pred, nonzero$volume_T)

```


### what is the proportion variation explained in the outcome of the testing data?
### i.e., what is 1-(SSerror/SStotal)
```{r, echo=TRUE}
actual <- valid$volume_T
predicted<-exp(valid$logVol_pred)
rsq <- 1-sum((actual-predicted)^2)/sum((actual-mean(actual))^2)
print(rsq)
```
